---
title: "day_8_data_viz"
author: "Julia Parish"
date: "8/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(janitor)


library(ggrepel)
library(patchwork)
library(ggrepel)
library(gghighlight)
library(paletteer)
library(ggExtra)
library(ggbeeswarm)
library(d3heatmap)

library(gapminder)

library(sf)
```

# Using the ‘space_launches.csv’ data (I’ll drop this into Slack), create a new .Rmd and create your worst possible graph by customizing (from the ggplot default) in at least **8 ways** - more encouraged - to make it as terrible as possible. - You can plot whatever variable(s) you choose, and make whatever graph type you want.
- Once you’re done, drop your worst graph (as a .png or .jpg) into the course Slack channel.

```{r}
space_launches <- read_csv(here("data","space_launches.csv"))

```
```{r}
space_plot <- ggplot(data = space_launches,
                     aes(x = mission, y = launch_year)) +
              geom_jitter(aes(color = launch_year),
                          show.legend = FALSE) +
              scale_color_gradientn(colors = c("chocolate4", "mediumspringgreen", "yellow3", "blueviolet")) +
  labs(title = "Missions Launched in Year",
       x = "Mission Title",
       y = "Launch Year") + 
  theme(
    plot.background = element_rect("lawngreen"),
    title = element_text(color = "orangered"),
    axis.text.x = element_text(color = "seagreen1", size = 4),
    axis.text.y = element_text(color = "mediumorchid", size = 16),
    axis.title = element_text(color = "hotpink", size = 18)
    )

space_plot  

ggsave(here::here("figs","myworst.png"),
       height = 5, width = 5)

```

### **Task 2: Wrangling practice with the SF Greenhouse Gas data**
- Download the San_Francisco_Communitywide_Greenhouse_Gas_Inventory.csv file (I’ll put this in Slack). Information about the data is available here.
- Read in the SF emissions dataset, then complete the following:

```{r}
sf_emissions <- read_csv(here("data", "San_Francisco_Communitywide_Greenhouse_Gas_Inventory.csv"))

```
# Create a summary table of total annual greenhouse gas emissions (only from Emissions_mtCO2e column) by sector_general
```{r}
e_annual <- sf_emissions %>% 
  group_by(Calendar_Year, Sector_General) %>% 
  summarize(annual_emissions = sum(Emissions_mtCO2e))

```
# Create a summary table of total annual greenhouse gas emissions (only from Emissions_mtCO2e column) by year and commodity type
```{r}
e_yr_type <- sf_emissions %>% 
  group_by(Calendar_Year, Commodity_Type) %>% 
  summarize(annual_emissions = sum(Emissions_mtCO2e))
```

# Create a subset of the data that only contains observations from 2005 on, only for observations where Sector_Detail2 contains the pattern “PG&E”, then limit to columns Calendar_Year, Sector_Detail2 and Emissions_mtCO2e.

```{r}
e_2005 <- sf_emissions %>% 
  filter(Calendar_Year > "2005") %>% 
  select(Calendar_Year, Sector_Detail2, Emissions_mtCO2e) %>% 
  filter(str_detect(Sector_Detail2, pattern = "PG&E"))

```
# **Task 3: Do your data viz best**
- Create a finalized graph (as perfect as you can make it) using the San Francisco GHG emissions dataset. You get to choose what you want to visualize. You can pick which variables & observations you’re interested in. You may want to do some wrangling / summarizing first.
- Once you’ve decided on what to plot, create the best graph that you can to communicate what’s going on with the data.
- Once you’re done, drop your best graph (as a .png or .jpg) into the course Slack channel.

```{r}
sf_emissions <- sf_emissions %>% 
  clean_names()

sf_e <- sf_emissions %>% 
  select(calendar_year, sector_general, emissions_mt_co2e) %>% 
  group_by(calendar_year, sector_general) %>% 
  summarize(emissions_mt_co2e = sum(emissions_mt_co2e)) %>% 
  filter(calendar_year != 2019) %>% 
  mutate(sector_general = fct_reorder(sector_general, emissions_mt_co2e))
  

```

```{r}
fancy_scientific <- function(l) {
  # turn in to character string in scientific notation
  l <- format(l, scientific = TRUE)
  # quote the part before the exponent to keep all the digits
  l <- gsub("^(.*)e", "'\\1'e", l)
  # turn the 'e+' into plotmath format
  l <- gsub("e", "%*%10^", l)
  # return this as an expression
  parse(text=l)
}
```


```{r}
sf_emissions_plot <- sf_e %>% 
  ggplot(aes(x = calendar_year, y = emissions_mt_co2e, 
             fill = sector_general)) +
  geom_area() +
  scale_x_continuous(breaks = c(1990, 2000, 2010, 2018),
                     limits = c(1990, 2018)) +
  scale_y_continuous(labels = fancy_scientific) +
  labs (title = "Greenhouse Gas Emissions in San Francisco",
        subtitle = "Source: San Francisco Communitywide Greenhouse Gas Inventory",
        x = "Year",
        y = "CO2 Emissions in MegaTons",
        fill = "Sectors") +
  theme(
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(size = 8), 
    panel.grid = element_line(color = "grey98") ,
    plot.background = element_rect (color = "floralwhite"),
    axis.text.x = element_text(color = "grey23"),
    axis.title = element_text(color = "grey10", size = 8, face = "bold"),
    panel.grid.minor.y = element_blank(),
    legend.position = "right",
    legend.title = element_text(face = "bold"),
    legend.text = element_text(size = 10)
    ) +
   scale_fill_paletteer_d(palette = "tvthemes::Stark") 
    
sf_emissions_plot

ggsave(here::here("figs","SFemissions_best.png"),
       height = 8, width = 8)
```

